---
title: "jazzPanda example"
date: "`r format(Sys.time(), '%d %b %Y')`"
vignette: >
    %\VignetteIndexEntry{jazzPanda example}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::rmarkdown}
output: rmarkdown::html_vignette
html_document:
    fig_caption: yes
    fig_retina: FALSE
    keep_md: FALSE
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE
)
```

# Introduction
Spatial transcriptomics allows the spatial profiling of complex tissue 
architectures. The spatial arrangement and interactions between cells can aid 
in understanding of complex functions and regulatory mechanisms in various 
tissue micro environments. Commercially available image-based spatial 
technologies such as Xenium, CosMX, and MERSCOPSE take advantage of 
fluorescence-based microscopy to quantify transcripts 
and focus on a pre-designed panel of genes. \n

One crucial step in the analysis of spatial transcriptomics data is cell 
type annotation. There are a number of ways to perform cell type annotation, 
and marker analysis is one of them. Marker gene analysis to identify genes 
highly expressed in each cluster compared to the remaining clusters. 
The identified marker genes are used to annotate clusters with cell types. 
Computationally tools originally developed for single cell data are used for 
spatial transcriptomics studies. However, those methods ignore the 
spatial information for the cells and gene. 
There is limited literature on developing marker gene detection methods 
that account for the spatial distribution of gene expression. 
jazzPanda provides two novel approaches to detect marker genes with transformed 
spatial information. Our first approach is based on correlation and the second 
linear modelling approach can account for technical noise and work for 
studies with multiple samples. \n

# jazzzPanda framework
We assume a marker gene will show a significant linear relationship with the 
target cluster over the tissue space. This suggests that the transcript 
detection of a marker gene will show similar patterns with the cells from
the cluster over tissue space. Given the cluster label of every cell, 
there are two steps to obtain marker genes for every cluster. We first compute 
spatial vectors from the spatial coordinates of the genes and the clusters. 
After that, we can measure the linear relationship between the genes and 
clusters based on spatial vectors. We develop two approaches for detecting 
genes that show strong linear relationship with the cluster. 

- **Step 1:** *Create spatial vectors for every gene and every cluster*  
    `get_vectors()` can be used to convert the transcript detection and the 
    cell centroids to spatial vectors. You can specify the tile shape and 
    length based on you dataset. The hex bins will generally take longer 
    than the square/rectangle bins to compute. 
    In practice, we find that we can choose the length tiles such that
    the average cell per tile
    ($\frac{\texttt{\#}\text{ cells per cluster} }{\texttt{\#}\text{ tiles}}$) 
    is close to one for each cluster. \n
    
- **Step 2:** *Detect marker genes*
    - Correlation approach: `compute_permp()`
    This approach can detect marker genes for one sample study. We calculate 
    a correlation coefficient between every pair of gene and cluster vector.
    We perform permutation testing to assess the statistical significance of 
    the calculated correlation and followed by multiple testing adjustment 
    to control the false discovery rate. We keep the genes with significant 
    adjusted p-value and large correlation as marker genes. 
    In practice, we recommend to calculate a correlation threshold value 
    for every cluster based on the data distribution. 
    During our analysis, we use 75\% quantile value of all correlations to 
    the given cluster as the cutoff and manage to keep meaningful marker genes. 
    
    - Linear modelling approach: `lasso_markers()`
    In this approach, we treat the gene vector as the response variable and 
    the cluster vectors as the explanatory variables. We select the important
    cluster vectors by lasso regularization first, and fit a linear model to 
    find the cluster that show minimum p-value and largest model coefficient.
    This approach can account for multi-sample studies and the technical 
    background noise (such as the non-specific binding).
    We recommend to set a model coefficient cutoff value based on your data. 
    A large cutoff value will result in fewer marker genes whereas 
    a small cutoff value will detect more marker genes.
    We find a cutoff value at 0.1 or 0.2 work well for our analysis. 
    Marker genes with less than 0.1 model coefficient are generally weak 
    markers.  
    Two tables will be returned from this approach:  
        1. We record the most significant cluster for each gene
        as `lasso_top_result`. This table provides unique marker genes for 
        every cluster. Genes whose top 
        cluster shows a model coefficient smaller than the specified cutoff 
        value will not be labeled as marker genes.  
        
        2. We record all the significant cluster for each gene as 
        `lasso_full_result`. This table can be used to investigate shared 
        marker genes for different clusters. 
        
# Example 
The dataset used in this vignette is a selected subset from two 
replicates of Xenium human breast cancer tissue Sample 1. We select 20 genes 
for package illustration. This subset was 
extracted from the raw dataset as described in the R script
located at `/inst/generate_vignette_data.R`. \n

This subset data is used for package illustration purpose only. 
The resulting marker genes may not be strong markers for annotating clusters. 
Please see the full analysis for this dataset for marker genes. 

```{r}
library(Seurat)
library(ggplot2)
library(jazzPanda)
library(data.table)
library(dplyr)
library(glmnet)
library(caret)
library(corrplot)
library(igraph)
library(ggraph)
library(ggrepel)
library(gridExtra)
library(utils)
library(spatstat)
library(tidyr)
library(ggpubr)
```

```{r}
# ggplot style
defined_theme = theme(strip.text = element_text(size = rel(1)),
                        strip.background = element_rect(fill = NA, 
                                                    colour = "black"),
                        axis.line=element_blank(),
                        axis.text.x=element_blank(),
                        axis.text.y=element_blank(),
                        axis.ticks=element_blank(),
                        axis.title.x=element_blank(),
                        axis.title.y=element_blank(),
                        legend.position="none",
                        panel.background=element_blank(),
                        panel.border=element_blank(),
                        panel.grid.major=element_blank(),
                        panel.grid.minor=element_blank(),
                        plot.background=element_blank())
```

## Load data 
```{r load subset of rep1}
data(rep1_sub, rep1_clusters, rep1_neg)
rep1_clusters$cluster=factor(rep1_clusters$cluster,
                            levels=paste("c",1:8, sep=""))
# record all real genes in the data
all_real_genes = unique(as.character(rep1_sub$trans_info$feature_name))
all_celltypes = unique(rep1_clusters[,c("anno","cluster")])
# nc_patterns = "^BLANK_|^NegControlProbe_|^NegControlCodeword_|^antisense_"
# nc_targets = grep(pattern =nc_patterns , x = all_genes,value = TRUE)
# all_real_genes = setdiff(all_genes, nc_targets)
```

The required input data structure for the main function `lasso_markers()` 
is a list of matrices. We will illustrate how to simply create the required 
input data from the raw output of different technologies in the following 
examples: 

### 10x Xenium data
The constructed xenium_data can be used in function `get_vectors()`.
The defined nc_lst can be used in function` create_genesets()`.
```{r xeium_example, eval=FALSE}

path <- 'path/to/data/directory'
mtx_name="cell_feature_matrix"
trans_name="transcripts.csv.gz"

transcript_info <- as.data.frame(fread(paste(path, trans_name,sep="")))
transcript_info$x <- as.numeric(transcript_info$x_location)
transcript_info$y <- as.numeric(transcript_info$y_location)
data <- Read10X(data.dir = paste(path,mtx_name, sep=""))
# count matrix 
cm <- as.matrix(data$`Gene Expression`)

# negative control targets
r_codeword <- as.matrix(data$`Negative Control Codeword`)
r_probe <- as.matrix(data$`Negative Control Probe`)

xenium_data = list(cm = cm, trans_info=transcript_info )

# -----------------------------------------------------------------------------
# cell label and coordinates 
clusters_info =  read.csv("./cells_meta.csv")
# the cluster label, sample label and coordinates for each cell 
cl_req_cols = c("x","y","cluster","sample")
clusters_df = clusters_info[,cl_req_cols]

# -----------------------------------------------------------------------------
# define negative control objects if available, need to repeat for each sample
# load coordinates for each detection, must contain following columns 
# - "x": x coordinate
# - "y": y coordinate 
# - "target": negative control names 
# - "sample": sample label 
probe_target = row.names(r_probe)
codeword=row.names(r_codeword)

nc_coords = read.csv("./negative_controls_detections.csv")
nc_names = unique(nc_coords$feature_name)
kpt_cols = c("x","y","feature_name","sample")
transcript_neg = transcript_info[transcript_info$feature_name %in% 
                                c(probe_target, codeword_target),
                                c("x_location","y_location","feature_name")]
colnames(transcript_neg)[1:3]= c("x","y","feature_name")
transcript_neg$sample = "sample1"
nc_lst = list(trans_info = transcript_neg, 
                probe=probe_target, 
                codeword=codeword)
```

### MERSCOPE data
The constructed merscope_data can be used in function `get_vectors()`.
The defined nc_lst can be used in function` create_genesets()`. 
```{r merscope example, eval=FALSE}
# count matrix
# rows as genes and columns as cells
# gene name as row names 
data_p = "path/to/merscope/output/data/"
cm = qread("./counts.qs")
# cell label and coordinates 
clusters_info =  read.csv("./cells_meta.csv")

# -----------------------------------------------------------------------------
# if transcript coordinates are available, need to repeat for each sample
# load transcript data, transcripts must contain follwing columns 
# - "x": x coordinate
# - "y": y coordinate 
# - "feature_name": gene name 
tr_req_cols = c("x", "y", "feature_name")
transcript_df =  read.csv("./transcripts.csv")
merscope_data = list(cm = cm, trans_info = transcript_df[,req_cols])
# the defined merscope_data can be used in function get_vectors() 

# the cluster label, sample label and coordinates for each cell 
cl_req_cols = c("x","y","cluster","sample")
clusters_df = clusters_info[,cl_req_cols]

# -----------------------------------------------------------------------------
# define negative control objects if available, need to repeat for each sample
# load coordinates for each detection, must contain following columns 
# - "x": x coordinate
# - "y": y coordinate 
# - "target": negative control names 
# - "sample": sample label 
nc_coords = read.csv("./negative_controls_detections.csv")
nc_names = unique(nc_coords$feature_name)
kpt_cols = c("x","y","feature_name","sample")
nc_lst = list(trans_info = nc_coords[,kpt_cols])

```

### CosMx data
The constructed cosmx_data can be used in function `get_vectors()`.
The defined nc_lst can be used in function` create_genesets()`. 
```{r cosmx example, eval=FALSE}
tiledbURI <- "file/path/to/TileDB/array/on/local/machine"

# read in SOMACollection
tiledb_scdataset <- tiledbsc::SOMACollection$new(uri = tiledbURI, 
                                                verbose = FALSE)
cm <- tiledb_scdataset$somas$RNA$X$members$counts$to_matrix(batch_mode = TRUE)

# the cluster label, sample label and coordinates for each cell 
clusters_info <- tiledb_scdataset$somas$RNA$obs$to_dataframe()

# -----------------------------------------------------------------------------
# if transcript coordinates are available, need to repeat for each sample
# load transcript data, transcripts must contain follwing columns 
# - "x": x coordinate
# - "y": y coordinate 
# - "feature_name": gene name 
req_cols = c("x","y","feature_name")
cosmx_data = list(cm = cm, trans_info = transcript_df[,c(req_cols)])
# 

# clusters and cell locations 
cl_req_cols = c("x","y","cluster","sample")
clusters_df = clusters_info[,cl_req_cols]

# -----------------------------------------------------------------------------
# if available
# define negative control objects if available
# need to repeat for each background category and each sample
# load coordinates for each detection, must contain following columns 
# - "x": x coordinate
# - "y": y coordinate 
# - "target": negative control names 
# - "sample": sample label 
nc_coords = read.csv("./negative_controls_detections.csv")
nc_names = unique(nc_coords$feature_name)
kpt_cols = c("x","y","feature_name","sample")
nc_lst = list(trans_info = nc_coords[,kpt_cols])

```

If you have a Seurat object, you can build the required object as follows.
The defined example_vectors_cm/example_vectors_tr can be passed 
to `lasso_markers` to identify marker genes. 

### from a Seurat object 
```{r eval=FALSE}
# cm is the count matrix
seu_obj =Seurat::CreateSeuratObject(counts = cm)
# make sure the clusters information contains column names: 
# cluster, x, y and sample
clusters_info = rep1_clusters
# make sure the transcript information contains column names: 
# feature_name, x, y
transcript_coords = rep1_sub$trans_info
data_example = list(cm=seu_obj@assays$RNA$counts,
                    trans_info =transcript_coords)

w_x =  c(min(floor(min(data$trans_info$x)), floor(min(clusters_info$x))), 
        max(ceiling(max(data$trans_info$x)), ceiling(max(clusters_info$x))))
w_y =  c(min(floor(min(data$trans_info$y)), floor(min(clusters_info$y))), 
        max(ceiling(max(data$trans_info$y)), ceiling(max(clusters_info$y))))

# build spatial vectors from transcript coordinates and cluster coordinates 
example_vectors_tr = get_vectors(data_lst= list(data_example),
                                cluster_info = clusters_info,
                                bin_type="square",
                                bin_param=c(10,10), 
                                all_genes = row.names(cm), 
                                w_x=w_x, w_y=w_y)
# -----------------------------------------------------------------------------
# build spatial vectors from count matrix and cluster coordinates 
# make sure the cluster information contains column names: 
# cluster, x, y, sample and cell_id
colnames(clusters_info)[5] = "cell_id"
example_vectors_cm = get_vectors(data_lst= NULL,
                                cm_lst=list(rep1=seu_obj@assays$RNA$counts),
                                cluster_info = clusters_info,
                                bin_type="square",
                                bin_param=c(10,10), 
                                all_genes = row.names(cm), 
                                w_x=w_x, w_y=w_y)
BiocCheck::.filter_data
```

## Visualise the clusters over the tissue space
We can plot the cells coordinates for each cluster of Replicate 1 subset
```{r rep clusters vis, fig.height=5,  fig.width=6,warning=FALSE}
p1<-ggplot(data = rep1_clusters,
        aes(x = x, y = y, color=cluster))+
        geom_point(position=position_jitterdodge(jitter.width=0,
                                        jitter.height=0), size=0.1)+
        scale_y_reverse()+
        theme_classic()+
        facet_wrap(~sample)+
        scale_color_manual(values = c("#FC8D62","#66C2A5" ,"#8DA0CB","#E78AC3",
                        "#A6D854","skyblue","purple3","#E5C498"))+
        guides(color=guide_legend(title="cluster", nrow = 2,
                        override.aes=list(alpha=1, size=2)))+
        
        theme(axis.text.x=element_blank(),
                axis.text.y=element_blank(),
                axis.ticks=element_blank(),
                axis.title.x=element_blank(),
                axis.title.y=element_blank(),
                panel.spacing = unit(0.5, "lines"),
                legend.position="none",
                strip.text = element_text(size = rel(1)))+
        xlab("")+
        ylab("")
p2<-ggplot(data = rep1_clusters,
        aes(x = x, y = y, color=cluster))+
        geom_point(position=position_jitterdodge(jitter.width=0, 
                                    jitter.height=0), size=0.1)+
        facet_wrap(~cluster, nrow = 2)+
        scale_y_reverse()+
        theme_classic()+
        scale_color_manual(values = c("#FC8D62","#66C2A5" ,"#8DA0CB","#E78AC3",
                        "#A6D854","skyblue","purple3","#E5C498"))+
        guides(color=guide_legend(title="cluster", nrow = 1,
        override.aes=list(alpha=1, size=4)))+
        theme(axis.text.x=element_blank(),
            axis.text.y=element_blank(),
            axis.ticks=element_blank(),
            axis.title.x=element_blank(),
            axis.title.y=element_blank(),
            panel.spacing = unit(0.5, "lines"), 
            legend.text = element_text(size=10),
            legend.position="none",
            legend.title = element_text(size=10),
            strip.text = element_text(size = rel(1)))+
        xlab("")+
        ylab("")

spacer <- patchwork::plot_spacer()
layout_design <- (p1 / spacer) | p2

layout_design <- layout_design + 
                    patchwork::plot_layout(widths = c(1, 4), heights = c(1, 1)) 

print(layout_design)
```

## Spatial vectors
We can visualize the spatial vectors for clusters and genes as follows. 
As an example for creating spatial vectors for genes, we plot the transcript
detections for the gene EPCAM over tissue space, along with the square and 
hex binning result. Similarly, we plot the cell coordinates in cluster c1, 
as well as the square and hex bin values over the space as an example. 
We can see that with the square and hex bins capture the key patterns of the
original coordinates. Hex bins can capture more details than square bins. 

### Example of gene vectors 
```{r plot spatial genes, fig.width=6, fig.height=4}

w_x =  c(min(floor(min(rep1_sub$trans_info$x)),
        floor(min(rep1_clusters$x))), 
        max(ceiling(max(rep1_sub$trans_info$x)),
        ceiling(max(rep1_clusters$x))))
w_y =  c(min(floor(min(rep1_sub$trans_info$y)),
        floor(min(rep1_clusters$y))), 
        max(ceiling(max(rep1_sub$trans_info$y)),
        ceiling(max(rep1_clusters$y))))

# plot transcript detection coordinates
selected_genes = rep1_sub$trans_info$feature_name == "EPCAM"
loc_mt = as.data.frame(rep1_sub$trans_info[selected_genes,
                        c("x","y","feature_name")]%>%distinct())
colnames(loc_mt)=c("x","y","feature_name")
layout(matrix(c(1, 2, 3), 1, 3, byrow = TRUE))
par(mar=c(5,3,6,3))
plot(loc_mt$x, loc_mt$y, main = "", xlab = "", ylab = "", 
        pch = 20, col = "maroon4", cex = 0.1,xaxt='n', yaxt='n')   
title(main = "EPCAM transcript detection", line = 3)
box() 

# plot square binning 
curr<-loc_mt[loc_mt[,"feature_name"]=="EPCAM",c("x","y")] %>% distinct()
curr_ppp <- ppp(curr$x,curr$y,w_x, w_y)
vec_quadrat <- quadratcount(curr_ppp, 10,10)
vec_its <- intensity(vec_quadrat, image=TRUE)
par(mar=c(0.01,1, 1, 2))
plot(vec_its, main = "")
title(main = "square binning", line = -2)

# plot hex binning 
w <- owin(xrange=w_x, yrange=w_y)
H <- hextess(W=w, 20)
bin_length <- length(H$tiles)
curr<-loc_mt[loc_mt[,"feature_name"]=="EPCAM",c("x","y")] %>% distinct()
curr_ppp <- ppp(curr$x,curr$y,w_x, w_y)
vec_quadrat <- quadratcount(curr_ppp, tess=H)
vec_its <- intensity(vec_quadrat, image=TRUE)
par(mar=c(0.1,1, 1, 2))
plot(vec_its, main = "")
title(main = "hex binning", line = -2)

```

### Example of cluster vectors 
```{r plot spatial clusters, fig.width=6, fig.height=4}
w_x =  c(min(floor(min(rep1_sub$trans_info$x)),
            floor(min(rep1_clusters$x))), 
        max(ceiling(max(rep1_sub$trans_info$x)),
            ceiling(max(rep1_clusters$x))))
w_y =  c(min(floor(min(rep1_sub$trans_info$y)),
            floor(min(rep1_clusters$y))), 
        max(ceiling(max(rep1_sub$trans_info$y)),
            ceiling(max(rep1_clusters$y))))

# plot cell coordinates
loc_mt = as.data.frame(rep1_clusters[rep1_clusters$cluster=="c1",
            c("x","y","cluster")])
colnames(loc_mt)=c("x","y","cluster")
layout(matrix(c(1, 2, 3), 1, 3, byrow = TRUE))
par(mar=c(5,3,6,3))
plot(loc_mt$x, loc_mt$y, main = "", xlab = "", ylab = "", 
        pch = 20, col = "maroon4", cex = 0.1,xaxt='n', yaxt='n')   
title(main = "cell coordinates in cluster c1", line = 3)
box() 

# plot square binning 
curr<-loc_mt[loc_mt[,"cluster"]=="c1", c("x","y")]%>%distinct()
curr_ppp <- ppp(curr$x,curr$y,w_x, w_y)
vec_quadrat <- quadratcount(curr_ppp, 10,10)
vec_its <- intensity(vec_quadrat, image=TRUE)
par(mar=c(0.1,1, 1, 2))
plot(vec_its, main = "")
title(main = "square binning", line = -2)

# plot hex binning 
w <- owin(xrange=w_x, yrange=w_y)
H <- hextess(W=w, 20)
bin_length <- length(H$tiles)
curr<-loc_mt[loc_mt[,"cluster"]=="c1",c("x","y")] %>%distinct()
curr_ppp <- ppp(curr$x,curr$y,w_x, w_y)
vec_quadrat <- quadratcount(curr_ppp, tess=H)
vec_its <- intensity(vec_quadrat, image=TRUE)
par(mar=c(0.1,1, 1, 2))
plot(vec_its, main = "")
title(main = "hex binning", line = -2)

```

The function `get_vectors()` can be used to create spatial vectors for
all the genes and clusters.
These spatial vectors may take the form of squares, rectangles, or hexagons 
specified by the `bin_type` parameter. 

### Spatial vectors for all genes and clusters
```{r create rep1 all vectors}
seed_number= 589

w_x =  c(min(floor(min(rep1_sub$trans_info$x)),
            floor(min(rep1_clusters$x))), 
        max(ceiling(max(rep1_sub$trans_info$x)),
            ceiling(max(rep1_clusters$x))))
w_y =  c(min(floor(min(rep1_sub$trans_info$y)),
            floor(min(rep1_clusters$y))), 
        max(ceiling(max(rep1_sub$trans_info$y)),
            ceiling(max(rep1_clusters$y))))

grid_length = 10
# get spatial vectors
rep1_sq10_vectors = get_vectors(data_lst= list(rep1_sub),
                                cluster_info = rep1_clusters,
                                bin_type="square",
                                bin_param=c(grid_length,grid_length), 
                                all_genes = all_real_genes , 
                                w_x=w_x, w_y=w_y)
```

The constructed spatial vectors can be used to quantify cluster-cluster and
gene-gene correlation.

#### Cluster-Cluster correlation
```{r fig.width=6, fig.height=6}
exp_ord = paste("c", 1:8, sep="")
rep1_sq10_vectors$cluster_mt = rep1_sq10_vectors$cluster_mt[,exp_ord]
cor_cluster_mt = cor(rep1_sq10_vectors$cluster_mt,
                rep1_sq10_vectors$cluster_mt, method = "pearson")
# Calculate pairwise correlations
cor_gene_mt = cor(rep1_sq10_vectors$gene_mt, rep1_sq10_vectors$gene_mt,
                    method = "pearson")

col <- grDevices::colorRampPalette(c("#4477AA", "#77AADD", 
                                    "#FFFFFF","#EE9988", "#BB4444"))

corrplot::corrplot(cor_cluster_mt, method="color", col=col(200), diag=TRUE,
                addCoef.col = "black",type="upper",
                tl.col="black", tl.srt=45, mar=c(0,0,5,0),sig.level = 0.05, 
                insig = "blank", 
                title = "cluster-cluster correlation (square bin = 40x40)"
                )
```

#### Gene-Cluster correlation
```{r gene_cluster_corr, warning=FALSE, message=FALSE,fig.width=6,fig.height=6}

cor_genecluster_mt = cor(x=rep1_sq10_vectors$gene_mt, 
                        y=rep1_sq10_vectors$cluster_mt, method = "pearson")

gg_correlation = as.data.frame(cbind(apply(cor_gene_mt, MARGIN=1, 
                                            FUN = mean, na.rm=TRUE),
                                        apply(cor_genecluster_mt, MARGIN=1, 
                                            FUN = mean, na.rm=TRUE)))
colnames(gg_correlation) = c("mean_correlation","mean_cluster")  
gg_correlation$gene=row.names(gg_correlation)

plot(ggplot(data = gg_correlation, 
    aes(x= mean_correlation, y=mean_cluster))+
        geom_point()+
        geom_text_repel(aes(label=gg_correlation$gene), size=1.8, hjust=1)+
    theme_bw()+
    theme(legend.title=element_blank(),
            axis.text.y = element_text(size=20),
            axis.text.x = element_text(size=20),
            axis.title.x=element_text(size=20),
            axis.title.y=element_text(size=20), 
            panel.spacing = unit(0.5, "lines"), 
            legend.position="none",
            legend.text=element_blank())+
    xlab("Average gene-gene correlation")+
    ylab("Average gene-cluster correlation"))

```

We can also construct a gene network based on the spatial vector for the
genes.

#### Gene network
```{r gene_graph, warning=FALSE, message=FALSE, fig.width=6, fig.height=6}

vector_graph= igraph::graph_from_adjacency_matrix(cor_gene_mt,
                                                mode = "undirected", 
                                                weighted = TRUE, 
                                                diag = FALSE)

vector_graph=igraph::simplify(igraph::delete.edges(vector_graph, 
                        E(vector_graph)[abs(E(vector_graph)$weight) <= 0.7]))

layout=igraph::layout_with_kk(vector_graph)

# Plot the graph
ggraph::ggraph(vector_graph, layout = layout) +
    geom_edge_link(aes(edge_alpha = weight), show.legend = FALSE) +
    geom_node_point(color = "lightblue", size = 5) +
    geom_node_text(aes(label = name), 
                    vjust = 1, hjust = 1,size=2,color="orange", repel = TRUE) 

```

## Linear relationship between markers and clusters

We assume that the relationship between a marker gene vector its cluster
spatial vector is linear. \n

Here are several genes and their annotation from the panel. \n

|         |                            |
|---------|----------------------------|
| Gene    | Annotation                 |
| ERBB2   | Breast cancer cells        |
| IL7R    | T cells                    |
| MZB1    | B cells                    |
| AQP1    | Endothelial                |
| LUM     | Fibroblasts                |

```{r linear vector to vector plot, fig.height=2, fig.width=7}
genes_lst = c("ERBB2","AQP1","LUM","IL7R","MZB1")

for (i_cluster in c("c1","c8","c3","c6","c7")){
    cluster_vector=rep1_sq10_vectors$cluster_mt[,i_cluster]
    
    data_vis=as.data.frame(cbind("cluster", cluster_vector,
                            rep1_sq10_vectors$gene_mt[, genes_lst]))
    
    colnames(data_vis)=c("cluster","cluster_vector",genes_lst)
    data_vis=reshape2::melt(data_vis,variable.name = "genes",
                            value.name = "gene_vector",
                            id= c("cluster","cluster_vector" ))
    data_vis$cluster_vector=as.numeric(data_vis$cluster_vector)
    data_vis$genes=factor(data_vis$genes)
    data_vis$gene_vector=as.numeric(data_vis$gene_vector)
    
    plot(ggplot(data = data_vis, 
            aes(x= cluster_vector, y=gene_vector))+
            geom_point(size=0.1)+
            facet_wrap(~genes,scales = "free_y", ncol=10)+
            theme_bw()+
            theme(legend.title=element_blank(),
                axis.text.y = element_text(size=6),
                axis.text.x = element_text(size=6,angle=0),
                axis.title.x=element_text(size=10),
                axis.title.y=element_text(size=10), 
                panel.spacing = unit(0.5, "lines"), 
                legend.position="none",
                legend.text=element_blank(),
                strip.text = element_text(size = rel(1)))+
                xlab(paste(i_cluster," - cluster vector", sep=""))+
                ylab("gene vector"))
}

```

## Scenario 1: one sample

A straightforward approach to identifying genes that exhibit a linear 
correlation with cluster vectors involves computing the Pearson correlation 
for each gene with every cluster. To assess the statistical significance 
of these correlations, the `compute_permp()` function can be used 
to perform permutation testing, generating a p-value for every pair of gene 
cluster and cluster vector. 

### Correlation-based method to detect marker genes 

```{r rep1 permutation}

w_x =  c(min(floor(min(rep1_sub$trans_info$x)),
            floor(min(rep1_clusters$x))), 
        max(ceiling(max(rep1_sub$trans_info$x)),
            ceiling(max(rep1_clusters$x))))
w_y =  c(min(floor(min(rep1_sub$trans_info$y)),
            floor(min(rep1_clusters$y))), 
        max(ceiling(max(rep1_sub$trans_info$y)),
            ceiling(max(rep1_clusters$y))))
set.seed(seed_number)
perm_p = compute_permp(data=rep1_sub,
                        cluster_info=rep1_clusters, 
                        perm.size=1000,
                        bin_type="square",
                        bin_param=c(10,10),
                        all_genes= all_real_genes,
                        correlation_method = "pearson", 
                        n.cores=1, 
                        correction_method="BH",
                        w_x=w_x ,
                        w_y=w_y)

# observed correlation for every pair of gene and cluster vector
obs_corr = as.data.frame(perm_p$obs.stat)
head(obs_corr)

# permutation adjusted p-value for every pair of gene and cluster vector
perm_res = as.data.frame(perm_p$perm.pval.adj)
head(perm_res)
```

#### Visualise top marker genes detected by correlation approach
Genes with a significant adjusted p-value are considered as marker genes for
the corresponding cluster. We can rank the marker genes by the observed 
correlationand plot the transcript detection coordinates for the top three
marker genes for every cluster. 
```{r rep1 permutation vis, fig.width=6, fig.height=3}
res_df_1000=as.data.frame(perm_p$perm.pval.adj)
res_df_1000$gene=row.names(res_df_1000)
cluster_names = unique(as.character(rep1_clusters$cluster))
for (cl in cluster_names){
    perm_sig = res_df_1000[res_df_1000[,cl]<0.05,]
    # define a cutoff value based on 75% quantile 
    obs_cutoff = quantile(obs_corr[, cl], 0.75)
    perm_cl=intersect(row.names(perm_res[perm_res[,cl]<0.05,]),
                        row.names(obs_corr[obs_corr[, cl]>obs_cutoff,]))
    inters=perm_cl
    rounded_val=signif(as.numeric(obs_corr[inters,cl]), digits = 3)
    inters_df = as.data.frame(cbind(gene=inters, value=rounded_val))
    inters_df$value = as.numeric(inters_df$value)
    inters_df=inters_df[order(inters_df$value, decreasing = TRUE),]
    inters_df = inters_df[1:min(nrow(inters_df),2),]
    inters_df$text= paste(inters_df$gene,inters_df$value,sep=": ")
    curr_genes = rep1_sub$trans_info$feature_name %in% inters_df$gene
    data_vis =rep1_sub$trans_info[curr_genes, c("x","y","feature_name")]
    data_vis$text = inters_df[match(data_vis$feature_name,inters_df$gene),
                                "text"]
    data_vis$text = factor(data_vis$text, levels=inters_df$text)
    p1<-ggplot(data = data_vis,
                aes(x = x, y = y))+
                geom_point(size=0.01,color="maroon4")+
                facet_wrap(~text,ncol=10, scales="free")+
                scale_y_reverse()+
                guides(fill = guide_colorbar(height= unit(5, "cm")))+
                defined_theme
    cl_pt<-ggplot(data = rep1_clusters[rep1_clusters$cluster==cl, ],
                    aes(x = x, y = y, color=cluster))+
                    geom_point(position=position_jitterdodge(jitter.width=0, 
                                            jitter.height=0), size=0.2)+
                    facet_wrap(~cluster)+
                    scale_y_reverse()+
                    theme_classic()+
                    scale_color_manual(values = "black")+
                    defined_theme
    lyt <- cl_pt | p1
    layout_design <- lyt + patchwork::plot_layout(widths = c(1,3)) 
    print(layout_design)
}
```

#### Visualise cluster vector and the top marker genes at spatial vector level
To check the linear relationship between the cluster vector and the marker 
gene vectors, we can plot the cluster vector on x-axis, and the marker gene 
vector on y-axis. The figure below shows the relationship between the 
cluster vector and the top marker gene vectors detected by correlation
approach. 
```{r vvplot_corr, fig.width=8, fig.height=6}

cluster_names = paste("c", 1:8, sep="")
plot_lst=list()
for (cl in cluster_names){
    perm_sig = res_df_1000[res_df_1000[,cl]<0.05,]
    curr_cell_type = all_celltypes[all_celltypes$cluster==cl,"anno"]
    obs_cutoff = quantile(obs_corr[, cl], 0.75)
    perm_cl=intersect(row.names(perm_res[perm_res[,cl]<0.05,]),
                        row.names(obs_corr[obs_corr[, cl]>obs_cutoff,]))
    inters=perm_cl
    rounded_val=signif(as.numeric(obs_corr[inters,cl]), digits = 3)
    inters_df = as.data.frame(cbind(gene=inters, value=rounded_val))
    inters_df$value = as.numeric(inters_df$value)
    inters_df=inters_df[order(inters_df$value, decreasing = TRUE),]
    inters_df$text= paste(inters_df$gene,inters_df$value,sep=": ")
    mk_gene = inters_df[1:min(2, nrow(inters_df)),"gene"]
    if (length(mk_gene > 0)){
        dff = as.data.frame(cbind(rep1_sq10_vectors$cluster_mt[,cl],
                                    rep1_sq10_vectors$gene_mt[,mk_gene]))
        colnames(dff) = c("cluster", mk_gene)
        
        dff$vector_id = c(1:(grid_length * grid_length))
        long_df <- dff %>% 
        pivot_longer(cols = -c(cluster, vector_id), names_to = "gene", 
                                values_to = "vector_count")
        long_df$gene = factor(long_df$gene, levels=mk_gene)
        
        p=ggplot(long_df, aes(x = cluster, y = vector_count )) +
            geom_point( size=0.01) +
            facet_wrap(~gene, scales = "free_y", nrow=1) +
            labs(x = paste("cluster vector ", curr_cell_type, sep=""), 
                    y = "marker gene vectors") +
            theme_minimal()+
            guides(color=guide_legend(nrow = 1,
                        override.aes=list(alpha=1, size=2)))+
            theme(panel.grid = element_blank(),legend.position = "none",
                    strip.text = element_text(size = rel(1)),
                    axis.line=element_blank(),
                    legend.title = element_blank(),
                    legend.key.size = unit(0.5, "cm"),
                    legend.text = element_text(size=10),
                    axis.text=element_blank(),
                    axis.ticks=element_blank(),
                    axis.title=element_text(size = 10),
                    panel.border =element_rect(colour = "black", 
                                                fill=NA, linewidth=0.5)
            )
        plot_lst[[cl]] = p
    }
}
combined_plot <- ggarrange(plotlist = plot_lst, 
                            ncol = 2, nrow = 4,
                            common.legend = FALSE, legend = "none")

combined_plot 

```
The other method to identify linearly correlated genes for each cluster is
to construct a linear model for each gene. We can use the `lasso_markers`
function to get the most relevant cluster label for every gene.

### Linear modeling approach to detect marker genes 
We can create spatial vectors for negative control genes and include
them as background noise "clusters".
```{r rep1 lasso_markers with background}

rep1_nc_vectors = create_genesets(data_lst=list("rep1"= rep1_neg),
                                    name_lst=list(probe=rep1_neg$probe, 
                                                codeword=rep1_neg$codeword), 
                                    bin_type="square",
                                    bin_param=c(10, 10), 
                                    w_x=w_x, w_y=w_y)

set.seed(seed_number)

rep1_lasso_with_nc = lasso_markers(gene_mt=rep1_sq10_vectors$gene_mt,
                                    cluster_mt = rep1_sq10_vectors$cluster_mt,
                                    sample_names=c("rep1"),
                                    keep_positive=TRUE, 
                                    coef_cutoff=0.2,
                                    background=rep1_nc_vectors)

rep1_top_df_nc = rep1_lasso_with_nc$lasso_top_result
# the top result table 
head(rep1_top_df_nc)

# the full result table 
rep1_full_df = rep1_lasso_with_nc$lasso_full_result
head(rep1_full_df)
```


#### Visualise top marker genes detected by linear modelling approach
We can rank the marker genes by its linear model coefficient to the cluster
ans plot the transcript detection coordinates for the top three marker genes
for every cluster. 
```{r lasso-rep1 vis, fig.width=6, fig.height=3}

cluster_names = paste("c", 1:8, sep="")
for (cl in setdiff(cluster_names,"NoSig")){
    inters=rep1_top_df_nc[rep1_top_df_nc$top_cluster==cl,"gene"]
    rounded_val=signif(as.numeric(rep1_top_df_nc[inters,"glm_coef"]),
                                    digits = 3)
    inters_df = as.data.frame(cbind(gene=inters, value=rounded_val))
    inters_df$value = as.numeric(inters_df$value)
    inters_df=inters_df[order(inters_df$value, decreasing = TRUE),]
    inters_df$text= paste(inters_df$gene,inters_df$value,sep=": ")
    
    if (length(inters > 0)){
        inters_df = inters_df[1:min(2, nrow(inters_df)),]
        inters = inters_df$gene
        iters_rep1= rep1_sub$trans_info$feature_name %in% inters
        vis_r1 =rep1_sub$trans_info[iters_rep1,
                                c("x","y","feature_name")]
        vis_r1$value = inters_df[match(vis_r1$feature_name,inters_df$gene),
                                "value"]
        #vis_r1=vis_r1[order(vis_r1$value,decreasing = TRUE),]
        vis_r1$text_label= paste(vis_r1$feature_name,
                                    vis_r1$value,sep=": ")
        vis_r1$text_label=factor(vis_r1$text_label, levels = inters_df$text)
        vis_r1$sample="rep1"
        p1<- ggplot(data = vis_r1,
                    aes(x = x, y = y))+ 
                    geom_point(size=0.01,color="maroon4")+
                    facet_wrap(~text_label,ncol=10, scales="free")+
                    scale_y_reverse()+
                    guides(fill = guide_colorbar(height= unit(5, "cm")))+
                    defined_theme
        cl_pt<-ggplot(data = rep1_clusters[rep1_clusters$cluster==cl, ],
                    aes(x = x, y = y, color=cluster))+
                    geom_point(position=position_jitterdodge(jitter.width=0, 
                                            jitter.height=0), size=0.2)+
                    facet_wrap(~cluster)+
                    scale_y_reverse()+
                    theme_classic()+
                    scale_color_manual(values = "black")+
                    defined_theme
        lyt <- cl_pt | p1
        layout_design <- lyt + patchwork::plot_layout(widths = c(1,3)) 
        print(layout_design)
}}
```


#### Visualise cluster vector and the top marker genes at spatial vector level
We can plot the cluster vector on x-axis, and the marker gene vectors (detected 
by the linear modelling approach) on y-axis to validate the linear relationship 
assumption between the cluster vector and the marker gene vectors. 
```{r vvplot_glm, fig.width=8, fig.height=6}

cluster_names = paste("c", 1:8, sep="")
plot_lst=list()
for (cl in cluster_names){
    curr_cell_type = all_celltypes[all_celltypes$cluster==cl,"anno"]
    inters=rep1_top_df_nc[rep1_top_df_nc$top_cluster==cl,"gene"]
    if (length(inters > 0)){
        rounded_val=signif(as.numeric(rep1_top_df_nc[inters,"glm_coef"]),
                                digits = 3)
        inters_df = as.data.frame(cbind(gene=inters, value=rounded_val))
        inters_df$value = as.numeric(inters_df$value)
        inters_df=inters_df[order(inters_df$value, decreasing = TRUE),]
        inters_df$text= paste(inters_df$gene,inters_df$value,sep=": ")
    
        inters_df = inters_df[1:min(2, nrow(inters_df)),]
        mk_gene = inters_df$gene

        dff = as.data.frame(cbind(rep1_sq10_vectors$cluster_mt[,cl],
                                rep1_sq10_vectors$gene_mt[,mk_gene]))
        colnames(dff) = c("cluster", mk_gene)
        dff$vector_id = c(1:(grid_length * grid_length))
        long_df <- dff %>% 
        pivot_longer(cols = -c(cluster, vector_id), names_to = "gene", 
                        values_to = "vector_count")
        long_df$gene = factor(long_df$gene, levels=mk_gene)
        p=ggplot(long_df, aes(x = cluster, y = vector_count )) +
            geom_point( size=0.01) +
            facet_wrap(~gene, scales = "free_y", nrow=1) +
            labs(x = paste("cluster vector ", curr_cell_type, sep=""), 
                    y = "marker gene vectors") +
            theme_minimal()+
            guides(color=guide_legend(nrow = 1,
                        override.aes=list(alpha=1, size=2)))+
            theme(panel.grid = element_blank(),legend.position = "none",
                    strip.text = element_text(size = rel(1)),
                    axis.line=element_blank(),
                    legend.title = element_blank(),
                    legend.key.size = unit(0.5, "cm"),
                    legend.text = element_text(size=10),
                    axis.text=element_blank(),
                    axis.ticks=element_blank(),
                    axis.title=element_text(size = 10),
                    panel.border =element_rect(colour = "black", 
                                                fill=NA, linewidth=0.5)
        )
        plot_lst[[cl]] = p
    }
}
combined_plot <- ggarrange(plotlist = plot_lst, 
                            ncol = 2, nrow = 4,
                            common.legend = FALSE, legend = "none")

combined_plot 

```

## Scenario 2: multiple samples
Load the replicate 2 from sample 1. 
```{r load rep2 data}
data(rep2_sub, rep2_clusters, rep2_neg)
rep2_clusters$cluster=factor(rep2_clusters$cluster,
                            levels=paste("c",1:8, sep=""))
rep1_clusters$cells = paste(row.names(rep1_clusters),"_1",sep="")
rep2_clusters$cells =paste(row.names(rep2_clusters),"_2",sep="")
rep_clusters = rbind(rep1_clusters,rep2_clusters)
rep_clusters$cluster=factor(rep_clusters$cluster,
                            levels=paste("c",1:8, sep=""))
table(rep_clusters$sample, rep_clusters$cluster)
```

#### Visualise the clusters 
We can plot the coordinates of cells for every cluster in every replicate 
```{r tworep cluster vis, fig.height=4,  fig.width=8, warning=FALSE}

ggplot(data = rep_clusters,
        aes(x = x, y = y, color=cluster))+
        geom_point(position=position_jitterdodge(jitter.width=0, 
                                                jitter.height=0),size=0.1)+
        facet_grid(sample~cluster)+
        scale_y_reverse()+
        theme_classic()+
        scale_color_manual(values = c("#FC8D62","#66C2A5" ,"#8DA0CB","#E78AC3",
                                "#A6D854","skyblue","purple3","#E5C498"))+
        guides(color=guide_legend(title="cluster", nrow = 1,
        override.aes=list(alpha=1, size=7)))+
        theme(
            axis.line=element_blank(),
            axis.text.x=element_blank(),
            axis.text.y=element_blank(),
            axis.ticks=element_blank(),
            axis.title.x=element_blank(),
            axis.title.y=element_blank(),
            panel.background=element_blank(),
            panel.border=element_blank(),
            panel.grid.major=element_blank(),
            panel.grid.minor=element_blank(),
            plot.background=element_blank(),
            legend.text = element_text(size=10),
            legend.position="none",
            legend.title = element_text(size=10),
            strip.text = element_text(size = rel(1)))+
        xlab("")+
        ylab("")
```

When we have multiple replicates in the dataset, we can find marker
genes by providing additional sample information as the input for the function
`lasso_markers`.

### Linear modeling approach to detect marker genes 
```{r tworep no background}
w_x =  c(min(floor(min(rep1_sub$trans_info$x)),
            floor(min(rep2_sub$trans_info$x)),
            floor(min(rep_clusters$x))), 
        max(ceiling(max(rep1_sub$trans_info$x)),
            ceiling(max(rep2_sub$trans_info$x)),
            ceiling(max(rep_clusters$x))))
w_y =  c(min(floor(min(rep1_sub$trans_info$y)),
            floor(min(rep2_sub$trans_info$y)),
            floor(min(rep_clusters$y))), 
        max(ceiling(max(rep1_sub$trans_info$y)),
            ceiling(max(rep2_sub$trans_info$y)),
            ceiling(max(rep_clusters$y))))
grid_length=10
# get spatial vectors
two_rep_vectors = get_vectors(data_lst= list("rep1"=rep1_sub, 
                                            "rep2" = rep2_sub),
                            cluster_info = rep_clusters, bin_type="square",
                            bin_param=c(grid_length, grid_length),
                            all_genes = all_real_genes , 
                            w_x=w_x, w_y=w_y)

two_rep_nc_vectors = create_genesets(data_lst=list("rep1" = rep1_neg, 
                                                    "rep2" = rep2_neg),
                                        name_lst=list(probe=rep1_neg$probe, 
                                                    codeword=rep1_neg$codeword),
                                        bin_type="square",
                                        bin_param=c(10,10), 
                                        w_x=w_x, w_y=w_y)
set.seed(seed_number)
two_rep_lasso_with_nc = lasso_markers(gene_mt=two_rep_vectors$gene_mt,
                                    cluster_mt = two_rep_vectors$cluster_mt,
                                    sample_names=c("rep1","rep2"),
                                    keep_positive=TRUE, 
                                    coef_cutoff=0.2,
                                    background=two_rep_nc_vectors,n_fold = 5)
tworep_res=two_rep_lasso_with_nc$lasso_top_result
tworep_res$celltype = rep_clusters[match(tworep_res$top_cluster,
                                            rep_clusters$cluster),"anno"]
table(tworep_res$top_cluster)
head(tworep_res)
```

#### Visualise the top marker genes for each cluster
```{r top3 marker gene, fig.height=6, fig.width=6}

for (cl in all_celltypes$anno){
    inters=tworep_res[tworep_res$celltype==cl,"gene"]
    rounded_val=signif(as.numeric(tworep_res[inters,"glm_coef"]),
                        digits = 3)
    inters_df = as.data.frame(cbind(gene=inters, value=rounded_val))
    inters_df$value = as.numeric(inters_df$value)
    inters_df=inters_df[order(inters_df$value, decreasing = TRUE),]
    inters_df$text= paste(inters_df$gene,inters_df$value,sep=": ")
    if (length(inters > 0)){
        inters_df = inters_df[1:min(2, nrow(inters_df)),]
        inters = inters_df$gene
        iters_rep1= rep1_sub$trans_info$feature_name %in% inters
        vis_r1 =rep1_sub$trans_info[iters_rep1,
                                c("x","y","feature_name")]
        vis_r1$value = inters_df[match(vis_r1$feature_name,inters_df$gene),
                                "value"]
        vis_r1=vis_r1[order(vis_r1$value,decreasing = TRUE),]
        vis_r1$text_label= paste(vis_r1$feature_name,
                                    vis_r1$value,sep=": ")
        vis_r1$text_label=factor(vis_r1$text_label)
        vis_r1$sample="rep1"
        iters_rep2= rep2_sub$trans_info$feature_name %in% inters
        vis_r2 =rep2_sub$trans_info[iters_rep2,
                                c("x","y","feature_name")]
        vis_r2$value = inters_df[match(vis_r2$feature_name,inters_df$gene),
                                "value"]
        vis_r2=vis_r2[order(vis_r2$value, decreasing = TRUE),]
        vis_r2$text_label= paste(vis_r2$feature_name,
                                vis_r2$value,sep=": ")
        vis_r2$text_label=factor(vis_r2$text_label)
        vis_r2$sample="rep2"
        p1<- ggplot(data = vis_r1,
                    aes(x = x, y = y))+ 
                    geom_point(size=0.01,color="maroon4")+
                    facet_grid(sample~text_label, scales="free")+
                    scale_y_reverse()+
                    guides(fill = guide_colorbar(height= unit(5, "cm")))+
                    defined_theme
        p2<- ggplot(data = vis_r2,
                    aes(x = x, y = y))+ 
                    geom_point(size=0.01,color="maroon4")+
                    facet_grid(sample~text_label,scales="free")+
                    scale_y_reverse()+
                    guides(fill = guide_colorbar(height= unit(5, "cm")))+
                    defined_theme
        
        cl_pt<-ggplot(data = rep_clusters[rep_clusters$anno==cl, ],
                    aes(x = x, y = y, color=cluster))+
                    geom_point(position=position_jitterdodge(jitter.width=0, 
                                            jitter.height=0), size=0.2)+
                    facet_grid(sample~cluster)+
                    scale_y_reverse()+
                    theme_classic()+
                    scale_color_manual(values = "black")+
                    defined_theme
        lyt <- cl_pt | (p1 / p2) 
        if (cl %in% c("c6","c7")){
            lyt <- cl_pt | ((p1 / p2) | patchwork::plot_spacer())   
        }
        layout_design <- lyt + patchwork::plot_layout(widths = c(1,3)) 

        print(layout_design)
}}
```

#### Visualise cluster vector and the top marker genes at spatial vector level
The figure below shows the relationship between the cluster vector and the 
top marker gene vectors detected by linear modelling approach by accounting 
for multiple samples and background noise 
```{r vvplot_twosample, fig.width=8, fig.height=6}
cluster_names = paste("c", 1:8, sep="")
plot_lst=list()
for (cl in cluster_names){
    inters=tworep_res[tworep_res$top_cluster==cl,"gene"]
    curr_cell_type = all_celltypes[all_celltypes$cluster==cl,"anno"]
    rounded_val=signif(as.numeric(tworep_res[inters,"glm_coef"]),
                            digits = 3)
    inters_df = as.data.frame(cbind(gene=inters, value=rounded_val))
    inters_df$value = as.numeric(inters_df$value)
    inters_df=inters_df[order(inters_df$value, decreasing = TRUE),]
    inters_df$text= paste(inters_df$gene,inters_df$value,sep=": ")
        
    mk_gene = inters_df[1:min(2, nrow(inters_df)),"gene"]
    if (length(inters > 0)){
        dff = as.data.frame(cbind(two_rep_vectors$cluster_mt[,cl],
                                    two_rep_vectors$gene_mt[,mk_gene]))
        colnames(dff) = c("cluster", mk_gene)
        total_tiles = grid_length * grid_length
        dff$vector_id = c(1:total_tiles)
        dff$sample= "Replicate1"
        dff[(total_tiles+1):(total_tiles*2),"sample"] = "Replicate2"
        dff$vector_id = c(1:total_tiles, 1:total_tiles)
        long_df <- dff %>% 
        pivot_longer(cols = -c(cluster, sample, vector_id), names_to = "gene", 
                        values_to = "vector_count")
        long_df$gene = factor(long_df$gene, levels=mk_gene)
        p=ggplot(long_df, 
                        aes(x = cluster, y = vector_count, color =sample)) +
            geom_point( size=0.01) +
            facet_wrap(~gene, scales = "free_y", nrow=1) +
            labs(x = paste("cluster vector ", curr_cell_type, sep=""), 
                    y = "marker gene vectors") +
            theme_minimal()+
            guides(color=guide_legend(nrow = 1,
                        override.aes=list(alpha=1, size=2)))+
            theme(panel.grid = element_blank(),legend.position = "bottom",
                    strip.text = element_text(size = rel(1)),
                    axis.line=element_blank(),
                    legend.title = element_blank(),
                    legend.key.size = unit(0.5, "cm"),
                    legend.text = element_text(size=10),
                    axis.text=element_blank(),
                    axis.ticks=element_blank(),
                    axis.title=element_text(size = 10),
                    panel.border =element_rect(colour = "black", 
                                                fill=NA, linewidth=0.5)
            )
        plot_lst[[cl]] = p
    }
}
combined_plot <- ggarrange(plotlist = plot_lst, 
                            ncol = 2, nrow = 4,
                            common.legend = TRUE, legend = "top")

combined_plot 
```


```{r session info}
sessionInfo()
```
